Memory dynamics for naturalistic experiences

### 2018-07-16 main.tex
new abstract

\section{Abstract}
{
The human memory system is adept at cataloging the rich dynamics of ongoing experience. However, traditional trial-based memory experiments cannot capture these dynamics, and therefore cannot be used to study them.  By constraining participants' experiences in an experiment to occur in temporally discrete trials (and often arranged in a randomized order) the temporal, contextual, and emotional structure of those in-lab experiences necessarily differ from the naturalistic experiences we encounter in everyday life.  Here we investigate how people verbally recall continuous videos by characterizing and relating the thematic dynamics, or ``trajectories,'' of the stimulus and participants' recalls.  Unlike trial-based studies of memory wherein participants attempt to recall the precise stimuli they encounter, naturalistic recall entails capturing the fundamental geometric components of the stimulus topic trajectory.  The precise words participants use to describe the stimulus, and the level of detail and number of distinct events they recall, vary considerably across participants.  Nevertheless, the majority of participants' recall narratives captured the fundamental ``shape'' of the original stimulus.  These findings provide a window into which aspects of naturalistic experiences must be preserved, and which might be more flexible, in considering whether and how those experiences are remembered.

>>>>

\section{Abstract}
{
The human memory system is adept at cataloging the rich dynamics of ongoing experience. A hallmark feature of human experiences is that they unfold in a structured and predictable manner, governed by the laws of the physical systems that produce them. However, this temporal structure is typically absent (or severly impoverished) in traditional trial-based memory experiments. As a consequence, our understanding of how the inherent temporal structure of our life experiences influences memorability and memory organization is limited. Using a novel computational modelling approach, we investigate how people verbally recall continuous videos by characterizing and relating the thematic dynamics, or ``trajectories,'' of the stimulus and participants' recalls. We found that despite large differences in the quantity of information recalled across participants, the overall shape of the stimulus was recapitulated by most of the participants. These findings provide a window into which aspects of naturalistic experiences must be preserved, and which might be more flexible, in considering whether and how those experiences are remembered.

% The human memory system is adept at cataloging the rich dynamics of ongoing experience. However, traditional trial-based memory experiments cannot capture these dynamics, and therefore cannot be used to study them.  By constraining participants' experiences in an experiment to occur in temporally discrete trials (and often arranged in a randomized order) the temporal, contextual, and emotional structure of those in-lab experiences necessarily differ from the naturalistic experiences we encounter in everyday life.  Here we investigate how people verbally recall continuous videos by characterizing and relating the thematic dynamics, or ``trajectories,'' of the stimulus and participants' recalls.  Unlike trial-based studies of memory wherein participants attempt to recall the precise stimuli they encounter, naturalistic recall entails capturing the fundamental geometric components of the stimulus topic trajectory.  The precise words participants use to describe the stimulus, and the level of detail and number of distinct events they recall, vary considerably across participants.  Nevertheless, the majority of participants' recall narratives captured the fundamental ``shape'' of the original stimulus.  These findings provide a window into which aspects of naturalistic experiences must be preserved, and which might be more flexible, in considering whether and how those experiences are remembered.
